---
title: "p8105_hw2_jl6321.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
```

## Problem 2

```{r mr_trash}
data_mr_trash = read_excel("Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>%
  janitor::clean_names()
data_mr_trash$sports_balls = as.integer(round(data_mr_trash$sports_balls)) 
```

```{r prof_trash}
data_prof_trash = read_excel("Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>%
  janitor::clean_names()
```

```{r join_dfs}
data_mr_trash$trash_wheel = "mr"
data_prof_trash$trash_wheel = "prof"
data_mr_trash$year = as.double(data_mr_trash$year)
data_joined_trash = full_join(data_mr_trash, data_prof_trash)
```

The number of observations in the resulting dataset is `r nrow(data_joined_trash)`. Some examples of key variables are dumpster, year, weight_tons, and sports_balls. The total weight of trash collected by Professor Trash Wheel is `r sum(data_prof_trash$weight_tons)` tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(filter(data_mr_trash, year == 2020)$sports_balls)`. 
